---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(AER)
library(truncreg)
library(margins)

#R Markdown files automatically set the wd to their directory
#Because of that the working directory should be the same as the directory of this file 
getwd() 

labsupp <- read.csv(file = 'mroz_wool.csv')
head(labsupp)
```

&nbsp;  

# OLS
```{r}
OLS <- lm(hours ~ kidslt6 + kidsge6 + age + educ + faminc + city + exper + expersq, data = labsupp)
summary(OLS)
```

&nbsp;  

# 1)
```{r}
tobit <- tobit(hours ~ kidslt6 + kidsge6 + age + educ + faminc + city + exper + expersq,
               data = labsupp)
summary(tobit)
tobittable <- coef(summary(tobit))
```

&nbsp;  

# 2)
```{r}
trunc <- truncreg(hours ~ kidslt6 + kidsge6 + age + educ + faminc + city + exper + expersq, data = labsupp)
summary(trunc)
trunctable <- coef(summary(trunc))
```
  
&nbsp;  

# 3)
Wage is highly endogenous to hours. Hours affect wages and vice versa. "bad control" problem. The regression equation can be interpreted as how family care affects labour supply of women of same age, education, experience and family income living in the same city.

&nbsp;  

# 4)
```{r}
income <- lm(hours ~ faminc, data = labsupp)
summary(income)

margins(income)
```

```{r}
tobitinc <- tobit(hours ~ faminc, data = labsupp)
summary(tobitinc)

```

The coefficient from the tobit regression is smaller than that from OLS when using all observations. 

```{r}
incomecond <- lm(hours ~ faminc, data = subset(labsupp, inlf == 1))
summary(incomecond)

margins(incomecond)
```



```{r}
tobitinccond <- tobit(hours ~ faminc, data = subset(labsupp, inlf == 1))
summary(tobitinccond)

```

When only considering those working, the two coefficients are identical (only the standard errors vary)
