---
title: '3.3 Markdown'
output: html_document
date: '2022-06-15'
---
# Load Dataset and required packages
```{r}
rm(list = ls())
# install.packages("tree")
library(tree)
# install.packages("ISLR")
library(ISLR)

carseats <- read.csv(file = 'carseats.csv')

# View(carseats)

```

# Exercise 1
Split the variables sales into a binary outcome variable with High = 1 if Sales > 8 and 0
otherwise.
```{r}
carseats$High <- ifelse(carseats$Sales > 8, 1, 0)

table(carseats$Sales, carseats$High)
```

# Exercise 2
Estimate an unpruned classification regression tree for the High/Low sales variable with
minimum size node of 12.

```{r}
seat_tree = tree(Sales ~ ., data = carseats, minsize = 12)
```
# Exercise 3
Plot the unpruned classification regression tree.

```{r}
plot(seat_tree)
text(seat_tree, pretty = 0)
title(main = "Unpruned Classification Tree")
```

# Exercise 4
Report for the unpruned classification regression tree the number of terminal nodes, the
residual deviance and the misclassification error rate.

```{r}
summary(seat_tree)
```

# Exercise 5
Split the total sample randomly into a training data set and a test data set with 200 obser-
vations each
 
```{r}
dim(carseats)
set.seed(2)
seat_idx = sample(1:nrow(carseats), 200)
seat_trn = carseats[seat_idx,]
seat_tst = carseats[-seat_idx,]

```

# Exercise 6
Compute an unpruned tree with minimum node size 12 for the training data.

```{r}
seat_tree = tree(Sales ~ ., data = seat_trn, minsize = 12)

```

# Exercise 7
Predict the tree for the training data and for the test data and compare the major summary
statistics

```{r}
seat_trn_pred = predict(seat_tree, seat_trn, type = "class")
seat_tst_pred = predict(seat_tree, seat_tst, type = "class")
```

# Exercise 8
Report the confusion matrices for the test and the training data and the corresponding accu-
racy measures

```{r}
# train confusion
table(predicted = seat_trn_pred, actual = seat_trn$Sales)
# test confusion
table(predicted = seat_tst_pred, actual = seat_tst$Sales)

accuracy = function(actual, predicted) {
  mean(actual == predicted)
}

# train acc
accuracy(predicted = seat_trn_pred, actual = seat_trn$Sales)
#test acc
accuracy(predicted = seat_trn_pred, actual = seat_tst$Sales)

```

# Exercise 9
Use a ten-fold cross-validation on the tree of the test set with number of terminal nodes equal
to 9 and 14.

```{r}
seat_tree_cv = cv.tree(seat_tree, FUN = prune.misclass)
min_idx = which.min(seat_tree_cv$dev)
seat_tree_cv$size[min_idx]
seat_tree_cv$size[9]
seat_tree_cv$size[14]

```


