---
title: '3.3 Markdown'
output: html_document
date: '2022-06-15'
---

Matr.Nr.: 
Michael 1187814,
Lucas 01/951052,
Nico 01/716631,

# Load Dataset and required packages
- skip first column (column number)
```{r}
rm(list = ls()) # clears everything
# install.packages("tree")
library(tree)

carseats <- read.csv(file = 'carseats.csv', colClasses = (c("NULL", replicate(11, NA))))

# View(carseats)

```

# Exercise 1
Split the variables sales into a binary outcome variable with High = 1 if Sales > 8 and 0
otherwise.
```{r}
carseats$HighSales <- as.factor(ifelse(carseats$Sales <= 8, "Low", "High"))

table(carseats$Sales, carseats$HighSales)
```

# Exercise 2
Estimate an unpruned classification regression tree for the High/Low sales variable with
minimum size node of 12.
- make character variables factor before

```{r}
carseats$Urban <- as.factor(carseats$Urban)
carseats$ShelveLoc <- factor(carseats$ShelveLoc)
carseats$US <- as.factor(carseats$US)

seat_tree = tree(HighSales ~ . - Sales, data = carseats, minsize = 12)
```
# Exercise 3
Plot the unpruned classification regression tree.

```{r}
plot(seat_tree)
text(seat_tree, pretty = 0, cex = 0.5)
title(main = "Unpruned Classification Tree")
```

# Exercise 4
Report for the unpruned classification regression tree the number of terminal nodes, the
residual deviance and the misclassification error rate.

```{r}
summary(seat_tree)
```

# Exercise 5
Split the total sample randomly into a training data set and a test data set with 200 obser-
vations each
 
```{r}
dim(carseats)
set.seed(2)
seat_idx = sample(1:nrow(carseats), 200)
seat_trn = carseats[seat_idx,]
seat_tst = carseats[-seat_idx,]

```

# Exercise 6
Compute an unpruned tree with minimum node size 12 for the training data.

```{r}
seat_tree_trn = tree(HighSales ~ . - Sales, data = seat_trn, minsize = 12)
plot(seat_tree_trn)
text(seat_tree_trn, pretty = 0, cex = 0.5)
title(main = "Unpruned Classification Tree (Training)")
summary(seat_tree_trn)
```

# Exercise 7
Predict the tree for the training data and for the test data and compare the major summary
statistics

```{r}
seat_trn_pred = predict(seat_tree_trn, seat_trn, type = "class")
seat_tst_pred = predict(seat_tree_trn, seat_tst, type = "class")
```

# Exercise 8
Report the confusion matrices for the test and the training data and the corresponding accu-
racy measures

```{r}
# train confusion
table(predicted = seat_trn_pred, actual = seat_trn$HighSales)
# test confusion
table(predicted = seat_tst_pred, actual = seat_tst$HighSales)

accuracy = function(actual, predicted) {
  mean(actual == predicted)
}

# train acc
accuracy(predicted = seat_trn_pred, actual = seat_trn$HighSales)
#test acc
accuracy(predicted = seat_tst_pred, actual = seat_tst$HighSales)

```

# Exercise 9
Use a ten-fold cross-validation on the tree of the test set with number of terminal nodes equal
to 9 and 14.

```{r}
# When setting a seed in a codebox in markdown, it does so only for given codebox,
# such that we have to set a seed again for reporducibility
set.seed(420)

# cross validation using classification error rate as indicator
seat_tree_cv <- cv.tree(seat_tree_trn, FUN = prune.misclass)
# cross validation results: size -> terminal nodes, dev -> missclassifications, k -> cost complexity parameter
seat_tree_cv$size
seat_tree_cv$dev
# plot the error rate 
plot(seat_tree_cv$size, seat_tree_cv$dev/nrow(seat_trn), type = "b")

# 9 nodes
seat_tree_trn_cv_9 = cv.tree(seat_tree_trn, FUN = prune.misclass, best = 9)
plot(seat_tree_trn_cv_9)
text(seat_tree_trn_cv_9, cex = 0.5)


seat_trn_cv_9_pred <- predict(seat_tree_trn_cv_9, seat_trn, type = "class")
seat_tst_cv_9_pred <- predict(seat_tree_trn_cv_9, seat_tst, type = "class")

# train acc
accuracy(predicted = seat_trn_cv_9_pred, actual = seat_trn$HighSales)
#test acc
accuracy(predicted = seat_tst_cv_9_pred, actual = seat_tst$HighSales)

# it is not possible to generate a tree with 14 nodes because of the minsize requirement we used prior
# a tree with 14 terminal nodes is not considered (like in the original example)
# the prune.missclass function defaults to using the next higher treesize (which is 17)
# 14 nodes
seat_tree_trn_cv_14 = cv.tree(seat_tree_trn, FUN = prune.misclass, best = 14)
plot(seat_tree_trn_cv_14)
text(seat_tree_trn_cv_14, cex = 0.5)

seat_trn_cv_14_pred <- predict(seat_tree_trn_cv_14, seat_trn, type = "class")
seat_tst_cv_14_pred <- predict(seat_tree_trn_cv_14, seat_tst, type = "class")

# train acc
accuracy(predicted = seat_trn_cv_14_pred, actual = seat_trn$HighSales)
#test acc
accuracy(predicted = seat_tst_cv_14_pred, actual = seat_tst$HighSales)



```

From Cross validation we know that the tree size of 12 seems to be as good as 9, let's try for fun
```{r}
# 12 nodes
seat_tree_trn_cv_12 = prune.misclass(seat_tree_trn, best = 12)
plot(seat_tree_trn_cv_12)
text(seat_tree_trn_cv_12, cex = 0.5)

seat_tree_trn_cv_12_pred <- predict(seat_tree_trn_cv_12, seat_trn, type = "class")
seat_tree_tst_cv_12_pred <- predict(seat_tree_trn_cv_12, seat_tst, type = "class")

table(predicted = seat_tree_trn_cv_12_pred, actual = seat_trn$HighSales)
accuracy(actual = seat_trn$HighSales, predicted = seat_tree_trn_cv_12_pred)

table(predicted = seat_tree_tst_cv_12_pred, actual = seat_tst$HighSales)
accuracy(actual = seat_tst$HighSales, predicted = seat_tree_tst_cv_12_pred)
```



