---
title: '3.3 Markdown'
output: html_document
date: '2022-06-15'
---
# Load Dataset and required packages
- skip first column (column number)
```{r}
rm(list = ls())
# install.packages("tree")
library(tree)
# install.packages("ISLR")
library(ISLR)

carseats <- read.csv(file = 'carseats.csv', colClasses = (c("NULL", replicate(11, NA))))

# View(carseats)

```

# Exercise 1
Split the variables sales into a binary outcome variable with High = 1 if Sales > 8 and 0
otherwise.
```{r}
carseats$HighSales <- as.factor(ifelse(carseats$Sales > 8, "Low", "High"))

# table(carseats$Sales, carseats$HighSales)
```

# Exercise 2
Estimate an unpruned classification regression tree for the High/Low sales variable with
minimum size node of 12.
- make character variables factor before

```{r}
carseats$Urban <- as.factor(carseats$Urban)
carseats$ShelveLoc <- factor(carseats$ShelveLoc)
carseats$US <- as.factor(carseats$US)

seat_tree = tree(HighSales ~ . - Sales, data = carseats, minsize = 12)
```
# Exercise 3
Plot the unpruned classification regression tree.

```{r}
plot(seat_tree)
text(seat_tree, pretty = 0)
title(main = "Unpruned Classification Tree")
```

# Exercise 4
Report for the unpruned classification regression tree the number of terminal nodes, the
residual deviance and the misclassification error rate.

```{r}
summary(seat_tree)
```

# Exercise 5
Split the total sample randomly into a training data set and a test data set with 200 obser-
vations each
 
```{r}
dim(carseats)
set.seed(2)
seat_idx = sample(1:nrow(carseats), 200)
seat_trn = carseats[seat_idx,]
seat_tst = carseats[-seat_idx,]

```

# Exercise 6
Compute an unpruned tree with minimum node size 12 for the training data.

```{r}
seat_tree_trn = tree(HighSales ~ . - Sales, data = seat_trn, minsize = 12)
plot(seat_tree_trn)
text(seat_tree_trn, pretty = 0)
title(main = "Unpruned Classification Tree (Training)")
summary(seat_tree_trn)
```

# Exercise 7
Predict the tree for the training data and for the test data and compare the major summary
statistics

```{r}
seat_trn_pred = predict(seat_tree_trn, seat_trn, type = "class")
seat_tst_pred = predict(seat_tree_trn, seat_tst, type = "class")
```

# Exercise 8
Report the confusion matrices for the test and the training data and the corresponding accu-
racy measures

```{r}
# train confusion
table(predicted = seat_trn_pred, actual = seat_trn$HighSales)
# test confusion
table(predicted = seat_tst_pred, actual = seat_tst$HighSales)

accuracy = function(actual, predicted) {
  mean(actual == predicted)
}

# train acc
accuracy(predicted = seat_trn_pred, actual = seat_trn$HighSales)
#test acc
accuracy(predicted = seat_tst_pred, actual = seat_tst$HighSales)

```

# Exercise 9
Use a ten-fold cross-validation on the tree of the test set with number of terminal nodes equal
to 9 and 14.

```{r}
# 9 nodes
seat_tree_trn_cv_9 = cv.tree(seat_tree_trn, FUN = prune.misclass, best = 9)
plot(seat_tree_trn_cv_9)
text(seat_tree_trn_cv_9)


seat_trn_cv_9_pred <- predict(seat_tree_trn_cv_9, seat_trn, type = "class")
seat_tst_cv_9_pred <- predict(seat_tree_trn_cv_9, seat_tst, type = "class")

# train acc
accuracy(predicted = seat_trn_cv_9_pred, actual = seat_trn$HighSales)
#test acc
accuracy(predicted = seat_tst_cv_9_pred, actual = seat_tst$HighSales)

# 14 nodes
seat_tree_trn_cv_14 = cv.tree(seat_tree_trn, FUN = prune.misclass, best = 14)
plot(seat_tree_trn_cv_14)
text(seat_tree_trn_cv_14)

seat_trn_cv_14_pred <- predict(seat_tree_trn_cv_14, seat_trn, type = "class")
seat_tst_cv_14_pred <- predict(seat_tree_trn_cv_14, seat_tst, type = "class")

# train acc
accuracy(predicted = seat_trn_cv_14_pred, actual = seat_trn$HighSales)
#test acc
accuracy(predicted = seat_tst_cv_14_pred, actual = seat_tst$HighSales)



```



